### 2014-11-10 doc/thesis/system.tex
Writing the examples from chapter 3 in the verbatim abstract syntax

>>>>

Função recursiva (another example?).

\begin{verbatim}
    rec f:(number) -> (number) =
      fun (x:number)
        if n == 0 then
          return 1
        else
          return n * f(n - 1)
    in f(5)
\end{verbatim}

Dynamic typing and static typing in the same code (another example?).

\begin{verbatim}
    local a:(number) -> (number) =
      fun (x:number)
        if n < 0 then return -n else return n
    in local d:(any, any) -> (number) =
         fun (x:any, y:any)
           skip ; return a(x - y)
       in d(1, 1)
\end{verbatim}

Another example:

\begin{verbatim}
    local x:any = 1 in x = x + 1
\end{verbatim}

This example is safe, but the following is not:

\begin{verbatim}
    local x:any, y:any = 1 in x = x + y
\end{verbatim}

Functions and multiple return.

%\begin{center}
%\begin{tabular}{llll}
%\multicolumn{4}{l}{$\mathbf{local} \; m:\Void \rightarrow \Number \times \String =$} \\
%& \multicolumn{3}{l}{$\mathbf{fun} \; ():\Number \times \String$} \\
%& & \multicolumn{2}{l}{$\mathbf{skip} \;;\; \mathbf{return} \; 2, ``foo"$} \\
%\multicolumn{4}{l}{$\mathbf{in}$} \\
%& \multicolumn{3}{l}{$\mathbf{local} \; s:\Number \times \Number \rightarrow \Number =$} \\
%& & \multicolumn{2}{l}{$\mathbf{fun} \; (x:\Number, y:\Number):\Number$} \\
%& & & \multicolumn{1}{l}{$\mathbf{skip} \;;\; \mathbf{return} \; x + y$} \\
%& \multicolumn{3}{l}{$\mathbf{in}$} \\
%& & \multicolumn{2}{l}{$\mathbf{local} \; x:\Number, y:\Number, z:\String = m(), m()$} \\
%& & \multicolumn{2}{l}{$\mathbf{in} \; s(m(), m())$}
%\end{tabular}
%\end{center}

\begin{verbatim}
  local m:() -> (number, string) =
    fun () skip ; return 2, "foo"
  in local s:(number, number) -> (number)
       fun (x:number, y:number) skip ; return x + y
     in local x:number, y:number, z:string = m(), m()
        in s(m(), m())
\end{verbatim}

Optional parameters.

\begin{verbatim}
    local m:(string, string|nil) -> (string) =
      fun (n:string, g:string|nil)
        g = g or "Hello" ; return g .. n
    in m("Lua")
\end{verbatim}

Overloading on the input parameter (another example?).

\begin{verbatim}
  local o:(string, string|integer) -> (string) =
    fun (a:string, b:string|integer)
      if type(b) == "string" then
        return a .. b
      else
        return string["rep"](a, b)
  in o("foo", 2) 
\end{verbatim}

Overloading on the return type (another example?).

\begin{verbatim}
  local d:(integer, integer) -> (integer, integer) | (nil, string) =
    fun (a:integer, b:integer):(integer, integer) | (nil, string)
      if b == 0 then
        return nil, "division by zero"
      else
        local r:integer = a \% b
        in local q:integer = (a - r) / b
           in skip ; return q, r
  in local q:X1, r:X2 = d(2, 2)
     in if q then
          print(a == b * q + r)
        else
          print("ERROR " .. r)
\end{verbatim}

Table type as hash.

\begin{verbatim}
    local t:{string:number|nil} = { ["foo"] = 1 } in
      local x:number|nil = t["foo"] in x = t["bar"]
\end{verbatim}

Table type as array.

\begin{verbatim}
    local d:{integer:string|nil} =
      { "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat" }
    in local x:number|nil = d[1] in x = d[8]
\end{verbatim}

Table type as record (first example).

\begin{verbatim}
    local p:{"firstname":string, "lastname":string} =
      { ["lastname"] = "Reed", ["firstname"] = "Lou" }
    in skip
\end{verbatim}

Table type as record (second example).

\begin{verbatim}
    local p:{"firstname":string,
             "middlename":string|nil,
             "lastname":string} =
      { ["firstname"] = "Lewis",
        ["middlename"] = "Allan",
        ["lastname"] = "Reed" }
    in skip
\end{verbatim}

Table type as record (third example).

\begin{verbatim}
    local p:{"firstname":string,
             "middlename":string|nil,
             "lastname":string} =
      { ["firstname"] = "Lou",
        ["lastname"] = "Reed" }
    in skip
\end{verbatim}

Refinement of table types (first example).

\begin{verbatim}
    local p:{} = {}
    in p["firstname"] <string> = "Lou" ;
       p["lastname"] <string> = "Reed"
\end{verbatim}

Refinement of table types (second example).

\begin{verbatim}
    local p:{} = {}
    in local b:{} = p
       in p["firstname"] <string> = "Lou" ;
          p["lastname"] <string> = "Reed"
\end{verbatim}

Modules (another example?).

\begin{verbatim}
    local m:{} = {}
    in local r:number = 3.14 / 180.0
       in m["deg"] <(number) -> (number)> =
          fun (x:number):(number)
            skip ; return x / r
        ; m["rad"] <(number) -> (number)> =
          fun (x:number):(number)
            skip ; return x * r
        ; m["pow"] <(number, number) -> (number)> =
          fun (x:number, y:number):(number)
            skip ; return x ^ y
\end{verbatim}

Objects.

\begin{verbatim}
    local s:{"x":number, "y":number} = { x = 0, y = 0 }
    in s["new"] <(self, number, number) -> (self)> =
       fun (self:self, x:number, y:number):(self)
         local s:{"x":number, "y":number} =
           setmetatable({}, { __index = self })
         in s["x"] = x ; s["y"] = y ; return s
     ; s["move"] <(self, number, number) -> ()> =
       fun (self:self, x:number, y:number):()
         self["x"] = self["x"] + x ;
         self["y"] = self["y"] + y 
\end{verbatim}
